### 2022-02-08 ms_body/manuscript.Rmd
Editing MS - add Fig. 7, add new refs

Species distribution models (SDM's) are becoming increasingly important tools to support and guide the development and implementation of environmental policies and management programs [@Addison2013; @Schuwirth2019]. The usefulness of SDM's is dependent on model performance, which is typically measured by its ability to correctly predict where a species is established and its capacity to distinguish between sites where the species is established versus a site where the species is absent (or assumed to be absent) [@Smith2021]. However, a number of authors have called in question the usefulness of SDM's, notably when models are transferred in space or time [@Yates2018; @Blasi2021]. Poor transferability may lead to erroneous model predictions that may have serious negative consequences for the application of the model, and resulting policies and management programmes (Soultan and Safi, 2017). Here, we report on the validation of a SDM for an insect, (*D. rubiformis*), native to Australia, that has been intentionally introduced as a biological control agent into South Africa to reduce the invasiveness of its host-plant, *A. mearnsii* (Impson et al., 2013). The the availability of independent testing data from South Africa, a separate continent to the data used to generate the models (Australia), makes this an ideal study system for assessing the transferability of SDM’s. This study represents one of only a handful of studies to validate SDM's using independent testing data [@Costa2010; @Rebelo2010; @West2016; @Smith2021].

@@

An additional concern with transferring MaxEnt models in space or time is the potential for models to significantly extrapolate predictions when the model is projected into non-analogous conditions to those under which the model was calibrated (Elith et al., 2010; Mesgaran et al., 2014). However, the applied use of SDM's in developing environmental policy, developing management programmes for invasive species and implementing conservation programmes often necessitate that models are transferred and predictions made, even when models are extrapolating (Elith et al., 2010). Here, we showed that *D. rubiformis* was able to establish at many sites (82%) in South Africa that were outside the range of climate space it occupies in its native range (i.e. extrapolation), which was not substantially different from establishment rates at sites in interpolation. While our MaxEnt models were able to make accurate predictions of *D. rubiformis* establishment in South Africa, even when models were extrapolating, this finding in no way discounts the importance of assessing model extrapolation when transferring models in space or time. A number of tools have been developed in recent years to assist SDM users with identifying model extrapolation, e.g. MESS [@Elith2010] and ExDet (Mesgaran et al., 2014), and which should be an integral component of any SDM study.

Given that the *D. rubiformis* MaxEnt models showed high predictive accuracy when transferred in space over South Africa and the statistically significant positive correlation between MaxEnt scores and *D. rubiformis* establishment, we wish to highlight how these models can be used in an applied setting. To visualise the potential geographic regions where *D. rubiformis* may establish in South Africa, we thresholded the MaxEnt map that optimised AICc where the probability of establishment was > 75% (i.e. MaxEnt score = 0.30), given in Fig. 7. The map demonstrates that the entire south and east coast regions of South Africa are climatically suitable for *D. rubiformis*. Mass-releasing and releasing weed biological control agents is a time-consuming and costly endeavor (Coetzee et al., 2021). As such, our models provide a statistically validated approach to optimising the redistribution of *D. rubiformis* in South Africa by prioritising releases at field sites where MaxEnt scores are > 0.30, and therefore where the probability of establishment was > 75%.

>>>>

Species distribution models (SDM's) are becoming increasingly important tools to support and guide the development and implementation of environmental policies and management programs [@Addison2013; @Schuwirth2019]. The usefulness of SDM's is dependent on model performance, which is typically measured by its ability to correctly predict where a species is established and its capacity to distinguish between sites where the species is established versus a site where the species is absent (or assumed to be absent) [@Smith2021]. However, a number of authors have called in question the usefulness of SDM's, notably when models are transferred in space or time [@Yates2018; @Blasi2021]. Poor transferability may lead to erroneous model predictions that may have serious negative consequences for the application of the model, and resulting policies and management programmes. Here, we report on the validation of a SDM for an insect, (*D. rubiformis*), native to Australia, that has been intentionally introduced as a biological control agent into South Africa to reduce the invasiveness of its host-plant, *A. mearnsii* [@Impson2013]. The the availability of independent testing data from South Africa, a separate continent to the data used to generate the models (Australia), makes this an ideal study system for assessing the transferability of SDM’s. This study represents one of only a handful of studies to validate SDM's using independent testing data [@Costa2010; @Rebelo2010; @West2016; @Smith2021].

@@

An additional concern with transferring MaxEnt models in space or time is the potential for models to significantly extrapolate predictions when the model is projected into non-analogous conditions to those under which the model was calibrated [@Elith2010; @Mesgaran2014]. However, the applied use of SDM's in developing environmental policy, developing management programmes for invasive species and implementing conservation programmes often necessitate that models are transferred and predictions made, even when models are extrapolating [@Elith2010]. Here, we showed that *D. rubiformis* was able to establish at many sites (82%) in South Africa that were outside the range of climate space it occupies in its native range (i.e. extrapolation), which was not substantially different from establishment rates at sites in interpolation. While our MaxEnt models were able to make accurate predictions of *D. rubiformis* establishment in South Africa, even when models were extrapolating, this finding in no way discounts the importance of assessing model extrapolation when transferring models in space or time. A number of tools have been developed in recent years to assist SDM users with identifying model extrapolation, e.g. MESS [@Elith2010] and ExDet [@Mesgaran2014], and which should be an integral component of any SDM study.

Given that the *D. rubiformis* MaxEnt models showed high predictive accuracy when transferred in space over South Africa and the statistically significant positive correlation between MaxEnt scores and *D. rubiformis* establishment, we wish to highlight how these models can be used in an applied setting. To visualise the potential geographic regions where *D. rubiformis* may establish in South Africa, we thresholded the MaxEnt map that optimised AICc where the probability of establishment was > 75% (i.e. MaxEnt score = 0.30), given in Fig. 7. The map demonstrates that the entire south and east coast regions of South Africa are climatically suitable for *D. rubiformis*. Mass-releasing and releasing weed biological control agents is a time-consuming and costly endeavor [@Hill2021]. As such, our models provide a statistically validated approach to optimising the redistribution of *D. rubiformis* in South Africa by prioritising releases at field sites where MaxEnt scores are > 0.30, and therefore where the probability of establishment was > 75%.

### 2022-02-08 ms_body/manuscript.tex
Editing MS - add Fig. 7, add new refs

Species distribution models (SDM's) are becoming increasingly important
tools to support and guide the development and implementation of
environmental policies and management programs (Addison et al., 2013;
Schuwirth et al., 2019). The usefulness of SDM's is dependent on model
performance, which is typically measured by its ability to correctly
predict where a species is established and its capacity to distinguish
between sites where the species is established versus a site where the
species is absent (or assumed to be absent) (Smith et al., 2021).
However, a number of authors have called in question the usefulness of
SDM's, notably when models are transferred in space or time (Blasi et
al., 2021; Yates et al., 2018). Poor transferability may lead to
erroneous model predictions that may have serious negative consequences
for the application of the model, and resulting policies and management
programmes (Soultan and Safi, 2017). Here, we report on the validation
of a SDM for an insect, (\emph{D. rubiformis}), native to Australia,
that has been intentionally introduced as a biological control agent
into South Africa to reduce the invasiveness of its host-plant, \emph{A.
mearnsii} (Impson et al., 2013). The the availability of independent
testing data from South Africa, a separate continent to the data used to
generate the models (Australia), makes this an ideal study system for
assessing the transferability of SDM's. This study represents one of
only a handful of studies to validate SDM's using independent testing
data (Costa et al., 2010; Rebelo and Jones, 2010; Smith et al., 2021;
West et al., 2016).

@@

Given that the \emph{D. rubiformis} MaxEnt models showed high predictive
accuracy when transferred in space over South Africa and the
statistically significant positive correlation between MaxEnt scores and
\emph{D. rubiformis} establishment, we wish to highlight how these
models can be used in an applied setting. To visualise the potential
geographic regions where \emph{D. rubiformis} may establish in South
Africa, we thresholded the MaxEnt map that optimised AICc where the
probability of establishment was \textgreater{} 50\% (i.e.~MaxEnt score
= 0.30), given in Fig. 9. The map demonstrates that the entire south and
east coast regions of South Africa are climatically suitable for
\emph{D. rubiformis}. Mass-releasing and releasing weed biological
control agents is a time-consuming and costly endeavor (Coetzee et al.,
2021). As such, our models provide a statistically validated approach to
optimising the redistribution of \emph{D. rubiformis} in South Africa by
prioritising releases at field sites where MaxEnt scores are
\textgreater{} 0.30, and therefore where the probability of
establishment was \textgreater{} 50\%.

@@

Species distribution models are becoming increasingly popular with
researchers, stakeholders, policy makers and land managers alike,
particularly for making predictions in novel geographic geographic
regions or forecasting species responses to climate change. Transfering
SDM's in space and/or time, however, has been met with justifiable
resilience and concerns over how well these models perform upon
transfer. In this study, we presented a case-study demonstrating that
MaxEnt SDM's were able to accurately predict the distribution of an
Australian insect when introduced into South Africa. Our study is one of
only a few studies to date validating MaxEnt models using spatially
independent data. However, the accuracy of the MaxEnt models and their
climatic suitability projections were dependent on the settings used to
calibrate the models. Our results demonstrate that MaxEnt can be useful
in predicting where a species may establish, even when projected into a
novel geographic area, but that users need to be aware of how model
settings and model extrapolation can influence model outputs and
usefulness. Demonstrating how SDM's can be useful, while communicating
uncertainty in their predictions and outputs, is essential for building
confidence in the use of SDM's (Blasi et al.~2020). We showed how
validating predictive models can be used to guide the implementation of
a biological control programme by optimising the redistribution and
release of \emph{D. rubiformis} in geographic areas that are
climatically suitable and where the probability of the insect
establishing is greater than 50\%.

@@

@@

@@

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{Default} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{HLQPT} & \cellcolor{gray!6}{0.89} & \cellcolor{gray!6}{0.04} & \cellcolor{gray!6}{0.18} & \cellcolor{gray!6}{564.46}\\
AUCtest & 6 & H & 0.89 & 0.02 & 0.09 & 15.56\\
\cellcolor{gray!6}{AUCdiff} & \cellcolor{gray!6}{4} & \cellcolor{gray!6}{HLQPT} & \cellcolor{gray!6}{0.89} & \cellcolor{gray!6}{0.02} & \cellcolor{gray!6}{0.13} & \cellcolor{gray!6}{485.91}\\
OR10 & 6 & HLQPT & 0.92 & 0.02 & 0.12 & 144.39\\
\cellcolor{gray!6}{AICc} & \cellcolor{gray!6}{5} & \cellcolor{gray!6}{H} & \cellcolor{gray!6}{0.90} & \cellcolor{gray!6}{0.02} & \cellcolor{gray!6}{0.09} & \cellcolor{gray!6}{0.00}\\*
\end{longtable}

>>>>

Species distribution models (SDM's) are becoming increasingly important
tools to support and guide the development and implementation of
environmental policies and management programs (Addison et al., 2013;
Schuwirth et al., 2019). The usefulness of SDM's is dependent on model
performance, which is typically measured by its ability to correctly
predict where a species is established and its capacity to distinguish
between sites where the species is established versus a site where the
species is absent (or assumed to be absent) (Smith et al., 2021).
However, a number of authors have called in question the usefulness of
SDM's, notably when models are transferred in space or time (Blasi et
al., 2021; Yates et al., 2018). Poor transferability may lead to
erroneous model predictions that may have serious negative consequences
for the application of the model, and resulting policies and management
programmes. Here, we report on the validation of a SDM for an insect,
(\emph{D. rubiformis}), native to Australia, that has been intentionally
introduced as a biological control agent into South Africa to reduce the
invasiveness of its host-plant, \emph{A. mearnsii} (Impson et al.,
2013). The the availability of independent testing data from South
Africa, a separate continent to the data used to generate the models
(Australia), makes this an ideal study system for assessing the
transferability of SDM's. This study represents one of only a handful of
studies to validate SDM's using independent testing data (Costa et al.,
2010; Rebelo and Jones, 2010; Smith et al., 2021; West et al., 2016).

@@

Given that the \emph{D. rubiformis} MaxEnt models showed high predictive
accuracy when transferred in space over South Africa and the
statistically significant positive correlation between MaxEnt scores and
\emph{D. rubiformis} establishment, we wish to highlight how these
models can be used in an applied setting. To visualise the potential
geographic regions where \emph{D. rubiformis} may establish in South
Africa, we thresholded the MaxEnt map that optimised AICc where the
probability of establishment was \textgreater{} 75\% (i.e.~MaxEnt score
= 0.30), given in Fig. 7. The map demonstrates that the entire south and
east coast regions of South Africa are climatically suitable for
\emph{D. rubiformis}. Mass-releasing and releasing weed biological
control agents is a time-consuming and costly endeavor (Hill et al.,
2021). As such, our models provide a statistically validated approach to
optimising the redistribution of \emph{D. rubiformis} in South Africa by
prioritising releases at field sites where MaxEnt scores are
\textgreater{} 0.30, and therefore where the probability of
establishment was \textgreater{} 75\%.

@@

Species distribution models are becoming increasingly popular with
researchers, stakeholders, policy makers and land managers alike,
particularly for making predictions in novel geographic geographic
regions or forecasting species responses to climate change. Transfering
SDM's in space and/or time, however, has been met with justifiable
resilience and concerns over how well these models perform upon
transfer. In this study, we presented a case-study demonstrating that
MaxEnt SDM's were able to accurately predict the distribution of an
Australian insect when introduced into South Africa. Our study is one of
only a few studies to date validating MaxEnt models using spatially
independent data. However, the accuracy of the MaxEnt models and their
climatic suitability projections were dependent on the settings used to
calibrate the models. Our results demonstrate that MaxEnt can be useful
in predicting where a species may establish, even when projected into a
novel geographic area, but that users need to be aware of how model
settings and model extrapolation can influence model outputs and
usefulness. Demonstrating how SDM's can be useful, while communicating
uncertainty in their predictions and outputs, is essential for building
confidence in the use of SDM's (Blasi et al.~2020). We showed how
validating predictive models can be used to guide the implementation of
a biological control programme by optimising the redistribution and
release of \emph{D. rubiformis} in geographic areas that are
climatically suitable and where the probability of the insect
establishing is greater than 75\%.

@@

\leavevmode\hypertarget{ref-Hill2021}{}%
Hill, M., Conlong, D., Zachariades, C., Coetzee, J., Paterson, I.,
Miller, B., Foxcroft, L., Van Der Westhuizen, L., 2021. The role of
mass-rearing in weed biological control projects in {South Africa}.
African Entomology 29, 1030--1044.

@@

\leavevmode\hypertarget{ref-Mesgaran2014}{}%
Mesgaran, M., Cousens, R., Webber, B., 2014. Here be dragons: {A} tool
for quantifying novelty due to covariate range and correlation change
when projecting species distribution models. Diversity and Distributions
20, 1147--1159. \url{https://doi.org/10.1111/ddi.12209}

@@

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{Default} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{HLQPT} & \cellcolor{gray!6}{0.89} & \cellcolor{gray!6}{0.04} & \cellcolor{gray!6}{0.18} & \cellcolor{gray!6}{564.46}\\
AUCtest & 6 & H & 0.89 & 0.02 & 0.09 & 15.56\\
\cellcolor{gray!6}{AUCdiff} & \cellcolor{gray!6}{4} & \cellcolor{gray!6}{HLQPT} & \cellcolor{gray!6}{0.89} & \cellcolor{gray!6}{0.02} & \cellcolor{gray!6}{0.13} & \cellcolor{gray!6}{485.91}\\
OR10 & 6 & HLQPT & 0.92 & 0.02 & 0.11 & 144.39\\
\cellcolor{gray!6}{AICc} & \cellcolor{gray!6}{5} & \cellcolor{gray!6}{H} & \cellcolor{gray!6}{0.90} & \cellcolor{gray!6}{0.02} & \cellcolor{gray!6}{0.09} & \cellcolor{gray!6}{0.00}\\*
\end{longtable}
